---
title: "Covid91 vaccine study Final2021"
author: "Michelle Coleman"
date: "April 24, 2021"
output: 
  html_document:
    number_sections: true
    toc: true

---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
FinalData= select(FinalData, ! is.numeric)
FinalData=purrr::map(FinalData, as.factor)
FinalData = data.frame(FinalData) 
FinalData$infected=relevel(FinalData$infected,"OK")
```



# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Methods

The first step is to isolate the subpopulations of Males, Females, DrugUsers, and LGBTQ.

```{r}
maledata <- filter(FinalData, sex=="guy")

```

# Males

## BarChart


```{r}
Males <- maledata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(Males,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Males,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
This Bar Chart appears to show that approximately 340 people were infected with Covid-19 57% of the males who had Covid were given the drug, and about 43% that were infected got the placebo.

## Numerial Summary

```{r}
table1 <- xtabs(~treatment + infected, data=maledata)
rowPerc(table1)
colPerc(table1)
print(table1)
```
The table shows that about 340 Males were infected with Covid-19 out of approximately 10,400 Males in the study above. The table also shows that 3.72% of Males who received the drug contracted Covid-19 and 2.55% of Males who received the placebo contracted Covid-19. Lastly the table shows that Males who got the vaccine were about 50% more to likely to contract Covid-19, this shows that the vaccine is not effective for Males.


### Inferential 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
The p-value is so small at 0.00006181 that is highly unlikely that the results were by chance.

The Null hypothesis is that the vaccine makes no difference in the Male population when it comes to contracting Covid-19.

The alternative hypothesis is that the vaccine in face does make a differnce in whether or not a Male contract Covid-19

Therefore, we reject the null hypothesis

The Fisher Exact test odds ratio in 0.68

Ultimately the vaccine is 40% effective in the Male population, which is absolutely ineffective.

# Females


```{r}
Femaledata <- filter(FinalData, sex=="gal")
Femaledata$sex <- as.factor(Femaledata$sex)
```

## Barcart

```{r}
Females <- Femaledata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(Females,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")


#Now for percentage plot

basicCC <- ggplot(Females,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This barchart shows that out of the approximately....Females who were infected with Covid-19, 32% were given the drug, and 66% were given placebo.


### Numerical Summaries  

```{r}
table2 <- xtabs(~treatment + infected, data=Femaledata)
rowPerc(table2)
colPerc(table2)
print(table2)
```

## Inferential  

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```
The value is 67% effective in Females.


# LGBTQ

```{r}
LGBTQdata <- filter(FinalData, LGBTQ=="gay")
LGBTQdata$sex <- as.factor(LGBTQdata$sex)
```

## Barcart

```{r}
LGBTQ <- LGBTQdata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(LGBTQ,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")


#Now for percentage plot

basicCC <- ggplot(LGBTQ,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```


### Numerical Summaries  

```{r}
table3 <- xtabs(~treatment + infected, data=LGBTQdata)
rowPerc(table3)
colPerc(table3)
print(table3)
```

## Inferential  

```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)

```

The vaccine is not effective in the LGBTQ population, it is only 29% effective.



# DrugUser

```{r}
DrugUserdata <- filter(FinalData, DrugUser=="yes")
DrugUserdata$DrugUser <- as.factor(DrugUserdata$DrugUser)
```


## Barcart

```{r}
DrugUser <- DrugUserdata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(DrugUser,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")


#Now for percentage plot

basicCC <- ggplot(DrugUser,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

### Numerical Summaries  

```{r}
table4 <- xtabs(~treatment + infected, data=DrugUserdata)
rowPerc(table4)
colPerc(table4)
print(table4)
```

## Inferential  

```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)

```

The vaccine is 60% effective in the Druggie population.


# Discussion & Conclusion

We looked at 4 populations, we found that the two populations that had the highest rate of effectiveness were the Females at 67% and the DrugUsers at 60%. Both the Male population and the LGBTQ populations showed that the vaccine was not beneficial to them and were ineffective in protecting them from contracting Covid-19, and Males had a 40% effectiveness, while the LGBTQ had 29% effectiveness, which is not at all effective.


